Supplementary Information

### 2020-11-26 doc/manuscript/MEE_manuscript.Rmd
move some content to SI, add site details appendix

# SUPPLEMENTARY INFO (this chunk must be repeated in MEE_manuscript_SI.Rmd)
#Appendices

DBH_reconstruction <- "S1"
climate_correction <- "S2"
traditional_comparison <- "S3"
changing_climate <- "S4"

DBH_reconstruction_title <- "Appendix S1. Methods for reconstruction of $DBH$"
climate_correction_title <- "Appendix S2. Methods for climate data evaluation and correction"
traditional_comparison_title <- "Appendix S3.  Methods for comparing climwin results with traditional methods"
changing_climate_title <- "Appendix S4.  Dealing with rapidly changing climate and tree growth"

@@

@@

For each year in the tree-ring records, we reconstructed $DBH$, as detailed in Appendix `r DBH_reconstruction`. 
In most cases, when a recent $DBH$ measurement was available, $DBH$ was reconstructed from the outside in. 
In cases where $DBH$ was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings [@applequist_simple_1958; @duncan_evaluation_1989], $DBH$ was reconstructed from the inside out. 
In either case, we used allometric equations between $DBH$ and bark thickness to account for changes in bark thickness as the tree grew (Appendix `r DBH_reconstruction`; Tables `r species_table`, `r bark_allometries`).

Once $DBH$ had been reconstructed, we used biomass allometries to estimate the corresponding aboveground biomass and diameter to area equation to get the corresponding basal area. We then calculated aboveground biomass growth increments ($\Delta AGB$) as [$AGB_{y+1}-AGB_y$] and basal area increment ($BAI$) as [$BAI_{y+1}-BAI_y$].
Biomass allometries for temperate and tropical sites were calculated using the R packages *allo-db* (Gonzalez-Akre et al. in prep) and *biomass* [@rejoumechain_biomass_2017], respectively.

Monthly climate data for 1901-2019 were obtained from CRU v.4.04 [@harris_updated_2014; @harris_version_2020], and in a few cases corrected based on more local records (Appendix `r climate_correction`). 
Variables considered here included average daily minimum, maximum, and mean temperatures ($T_{min}$, $T_{max}$, $T_{mean}$, respectively); precipitation ($PPT$); and, when deemed reliable (Appendix `r climate_correction`), potential evapotranspiration ($PET$) and precipitation day frequency ($PDF$).
*For BCI, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on BCI starting in 1929 [@paton_barro_2019].*
All ForestGEO climate records used here are archived in the ForestGEO Climate Data Portal, *v1.0-alpha* [@anderson-teixeira_forestgeoclimate_2020].

>>>>

# SUPPLEMENTARY INFO (this chunk must be repeated in MEE_manuscript_SI.Rmd)
#Appendices
site_details_appendix <- "S1"
DBH_reconstruction <- "S2"
climate_correction <- "S3"
traditional_comparison <- "S4"
changing_climate <- "S5"

site_details_appendix_title<- "Appendix S1. Methods for reconstruction of $DBH$"
DBH_reconstruction_title <- "Appendix S2. Methods for reconstruction of $DBH$"
climate_correction_title <- "Appendix S3. Methods for climate data evaluation and correction"
traditional_comparison_title <- "Appendix S4.  Methods for comparing climwin results with traditional methods"
changing_climate_title <- "Appendix S5.  Dealing with rapidly changing climate and tree growth"

@@

**Characterizing how tree growth is responding to global change (slowly-changing environmental drivers) is very challenging and uncertain.**

@@

For each year in the tree-ring records, we reconstructed $DBH$, as detailed in Appendix `r DBH_reconstruction`. 
We used allometric equations between $DBH$ and bark thickness to account for changes in bark thickness as the tree grew (Appendix `r DBH_reconstruction`; Tables `r species_table`, `r bark_allometries`).

Once $DBH$ had been reconstructed, we used biomass allometries to estimate the corresponding aboveground biomass and diameter to area equation to get the corresponding basal area. We then calculated aboveground biomass growth increments ($\Delta AGB$) as [$AGB_{y+1}-AGB_y$] and basal area increment ($BAI$) as [$BAI_{y+1}-BAI_y$].
Biomass allometries for temperate and tropical sites were calculated using the R packages *allo-db* (Gonzalez-Akre et al. in review) and *biomass* [@rejoumechain_biomass_2017], respectively.

Monthly climate data for 1901-2019 were obtained from CRU v.4.04 [@harris_updated_2014; @harris_version_2020], and in a few cases corrected based on more local records (Appendix `r climate_correction`). 
Variables considered here included average daily minimum, maximum, and mean temperatures ($T_{min}$, $T_{max}$, $T_{mean}$, respectively); precipitation ($PPT$); and, when deemed reliable (Appendix `r climate_correction`), potential evapotranspiration ($PET$) and precipitation day frequency ($PDF$).
All ForestGEO climate records used here are archived in the ForestGEO Climate Data Portal, *v1.0-alpha* [@anderson-teixeira_forestgeoclimate_2020].

### 2020-11-26 doc/manuscript/MEE_manuscript_SI.Rmd
move some content to SI, add site details appendix

#Appendices
DBH_reconstruction <- "S1"
climate_correction <- "S2"
traditional_comparison <- "S3"
changing_climate <- "S4"

DBH_reconstruction_title <- "Appendix S1. Methods for reconstruction of $DBH$"
climate_correction_title <- "Appendix S2. Methods for climate data evaluation and correction"
traditional_comparison_title <- "Appendix S3.  Methods for comparing climwin results with traditional methods"
changing_climate_title <- "Appendix S4.  Dealing with rapidly changing climate and tree growth"

@@

@@

@@

>>>>

#Appendices
site_details_appendix <- "S1"
DBH_reconstruction <- "S2"
climate_correction <- "S3"
traditional_comparison <- "S4"
changing_climate <- "S5"

site_details_appendix_title<- "Appendix S1. Methods for reconstruction of $DBH$"
DBH_reconstruction_title <- "Appendix S2. Methods for reconstruction of $DBH$"
climate_correction_title <- "Appendix S3. Methods for climate data evaluation and correction"
traditional_comparison_title <- "Appendix S4.  Methods for comparing climwin results with traditional methods"
changing_climate_title <- "Appendix S5.  Dealing with rapidly changing climate and tree growth"

@@

## `r site_details_appendix_title`

*(include descriptions of stand history, global change dynamics)*

### Barro Colorado Island, Panama

@@

In most cases, when a recent $DBH$ measurement was available, $DBH$ was reconstructed from the outside in. 
In cases where $DBH$ was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings [@applequist_simple_1958; @duncan_evaluation_1989], $DBH$ was reconstructed from the inside out.

@@

*For BCI, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on BCI starting in 1929 [@paton_barro_2019].*
